---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# measoshapes

The goal of measoshapes is to provide regionalization boundaries for MEASO. 

There are several in-built data sets currently at **version 05**. See below where each is explained. 

## Installation

To install measoshapes from Github use

```R
## install.packages("remotes")
remotes::install_github("AustralianAntarcticDivision/measoshapes")
```


## Data sets

There is no code in the measoshapes package, but the package records the creation of
each shapes layer using standard R mechanisms. 

There are five datasets, and for normal usage, calculating areas, and making maps we would use the first three. The `_coastline` forms have the continent of Antarctica cut out of the shapes. 


* `measonames` a dataframe of the `name`, `sector`, and `zone`
* `measo_regions05_ll_coastline` the polygons with the continent of Antarctica cut out
* `measo_regions05_coastline` the polygons (in polar) with the continent of Antarctica cut out

Two other forms in polar and longitude/latitude form are use for model overlays (for technical reasons). 

* `measo_regions05` the polygons of each combination of sector and zone
* `measo_regions05_ll` the polygons in longitude/latitude form


To make a map with these we can 

```{r measo-map}
library(measoshapes)
plot(measo_regions05_ll_coastline, reset = FALSE)
maps::map(add = TRUE, col = "grey", fill = FALSE)
```




To export to shapefile: 

```{r exportshp, eval = FALSE}
st_write(measo_regions05_ll_coastline, "myfile.shp", driver = "SHP")
```


---

Please note that the 'measoshapes' project is released with a
  [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.

