---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 7, fig.height = 7
)
library(dplyr)
library(sf)

```
# measoshapes

The goal of measoshapes is to provide regionalization boundaries for MEASO. 

There are several in-built data sets currently at **version 05**. See below where each is explained. 

## Installation

To install measoshapes from Github use

```R
## install.packages("remotes")
remotes::install_github("AustralianAntarcticDivision/measoshapes")
```

## Map

To map the regions with base plotting (with the [sf package](https://r-spatial.github.io/sf)). 

First join the names and properties to the geometry, we also group the features so that the seam at the anti-meridian is
removed. 

```{r join}
library(measoshapes)
library(dplyr)
library(sf)

measo <- measo_regions05 %>% group_by(name) %>% summarize() %>% 
  inner_join(measo_names)

```

Now, plot the geometry with the offical colours and make sure no 'reset' so that we can add other layers. 

```{r plot}
plot(st_geometry(measo), col = measo$fill, reset = FALSE, border = NA)
```

A very simple coastline data is added. 

```{r coast}
coast <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>% dplyr::filter(sovereignt == "Antarctica")
coast <- st_transform(coast, 
                      st_crs(measo))

plot(st_geometry(measo), col = measo$fill, reset = FALSE, border = NA)
plot(st_geometry(coast), col = "#808080", add = TRUE, border = NA)


```

## Map with ggplot2

A ggplot2 example. To use the literal fill colours we need `scale_fill_identity()`, this is ggplot's "straight-through" mechanism. 


```{r}

library(ggplot2)
ggplot(measo, aes(fill = fill)) + geom_sf(colour = NA) + scale_fill_identity() + geom_sf(data = coast, aes(fill = NULL))

```

## Map with SOmap

The [SOmap package](https://github.com/AustralianAntarcticDivision/SOmap) provides some simpler ways of creating and adding to polar maps. Once a map is set up it knows what projection it is in, so we can forget about having to transform the data. 

We can add arbitrary data at longitude/latitudes pairs and SOplot knows what to do. 

```{r SOmap}
## atm we need a special branch (2020-02-20)
#remotes::install_github("AustralianAntarcticDivision/SOmap@dev-0.6")
library(SOmap)
SOmap(trim = -32) ## set up a polar map
## set overlay to half-transparent
SOplot(st_geometry(measo), border = NA, col = scales::alpha(measo$fill, 0.75))  ## no need to worry about the projection

ll <- cbind(lon = c(147, 100, -10, -80), 
            lat = c(-42, -50, -70, -50))
SOplot(ll, pch = 1:4, col = viridis::viridis(4), cex = 3, lwd = 4)

## other objects work too (sf, raster, sp)
SOplot(SOmap_data$fronts_park)
SOplot(coast, col = "#808080")

```

## Data sets

There is no code in the measoshapes package, but the package records the creation of
each shapes layer using standard R mechanisms. 

There are five datasets, and for normal usage, calculating areas, and making maps we would use the first three. The `_coastline` forms have the continent of Antarctica cut out of the shapes. 


* `measonames` a dataframe of the `name`, `sector`, and `zone`
* `measo_regions05_ll_coastline` the polygons with the continent of Antarctica cut out
* `measo_regions05_coastline` the polygons (in polar) with the continent of Antarctica cut out

Two other forms in polar and longitude/latitude form are use for model overlays (for technical reasons). 

* `measo_regions05` the polygons of each combination of sector and zone
* `measo_regions05_ll` the polygons in longitude/latitude form


To make a map with these we can use the following code

```{r measo-map}
library(measoshapes)
plot(measo_regions05_ll_coastline, reset = FALSE)
maps::map(add = TRUE, col = "grey", fill = FALSE)
```




To export to shapefile: 


```{r exportshp1, eval = FALSE}
st_write(measo_regions05_ll_coastline, "myfile.shp", driver = "SHP")
```


Here we explore the names in more detail. The northern regions are assigned codes that include "T" for **temperate**, but they aren't supposed to be part of MEASO. They share a sector but have no assigned zone. From `colour_values` on zone they get black. 

```{r dummy}
par(mar = rep(0.2, 4))
plot(st_geometry(measo_regions05_ll), reset = FALSE, 
     col = colourvalues::colour_values(measo_names$sector, alpha = 0.5))
cds <- st_coordinates(st_centroid(measo_regions05_ll))
text(cds, lab = measo_names$name, cex = 0.7)
par(xpd = NA)
text(0, -20, "name")
maps::map(add = TRUE)
par(xpd = TRUE)

plot(st_geometry(measo_regions05_ll), reset = FALSE, 
     col = colourvalues::colour_values(measo_names$zone, alpha = 0.5))
cds <- st_coordinates(st_centroid(measo_regions05_ll))
text(cds, lab = measo_names$zone, cex = 0.7)
par(xpd = NA)
text(0, -20, "zone")
maps::map(add = TRUE)
par(xpd = TRUE)


plot(st_geometry(measo_regions05_ll), reset = FALSE, 
     col = colourvalues::colour_values(measo_names$zone, alpha = 0.5))
cds <- st_coordinates(st_centroid(measo_regions05_ll))
text(cds, lab = measo_names$sector, cex = 0.7)
par(xpd = NA)
text(0, -20, "sector")
maps::map(add = TRUE)
par(xpd = TRUE)

```

---

Please note that the 'measoshapes' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.

